---
title: "Machine_Learning_Proj_Coursera"
author: "Edayadulla M"
date: "8/24/2020"
output:
  html_document: default
  pdf_document: default
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```
# Project Objective
### Main Goal of this project is to understand , analyse and predict the manner in which     people did the exercise using the "Weight Lifting Exercise Dataset". 
### "classe" variable in the training dataset is the outcome we would like to predict.
## Below are the list of steps being done. 
### 1. Loading of Train and test datasets.
### 2. Quick Exploratory data analysis on the Dataset.
### 3. Data Clean-up. Assumption taken to disgard the predictor variables which are 
### having 'NA' values > 70% of the time.
### 4. Removed User_id and TimeStamp related variables.
### 5. Split clean dataset into Training & Testing datasets.
### 6. Using various models to train the "training" dataset.
### 7. Review the model results to choose most approprite one.
### 8. Predict the 20 input values provided in the test dataset for quiz
## Loading of all required libraries to execute this project
```{r 1. Library Load}
library(dplyr)
library(caret)
library(corrplot)
library(mltools)
library(e1071)
library(ggplot2)
library(randomForest)
source("http://www.sthda.com/upload/rquery_cormat.r")
require("corrplot")
library(rquery)
```
## Loading of Test & train datasets and Pre-processing 
```{r 2. Data down load and clean-up, echo=FALSE}
#plot(pressure)
train_data <- read.csv("C:/Users/IQBALE/R_Projects/pml-training.csv")
test_data <- read.csv("C:/Users/IQBALE/R_Projects/pml-testing.csv")
train_data <- train_data %>% mutate_all(na_if,"")
cols <- data.frame(sapply(train_data, function(x) sum(is.na(x))))
cols <- cbind(variable = row.names(cols), cols)
select_cols <- list() 
j = 1;
for ( i in 1:dim(cols)[1] )
   if ( as.integer(cols[i,2]) > (dim(train_data)[1] * 0.7) ){
      select_cols[j] <- as.character(cols[i,1])
      j = j+1
   }
train_clean1 <- train_data[,!(names(train_data) %in% select_cols)]
test_clean1 <- test_data[,!(names(test_data) %in% select_cols)]
remove_col <- c("X","new_window","user_name","raw_timestamp_part_1", "raw_timestamp_part_2","cvtd_timestamp","problem_id")

train_clean  <- train_clean1[,!(names(train_clean1) %in% remove_col)]
test_clean   <- test_clean1[,!(names(test_clean1) %in% remove_col)]
```
##  EXploratory Data Analysis
``` {r 2A.Co-relation b/w predictor variables , echo=FALSE }
rquery.cormat(train_clean[,1:20])
```           
##  Training and Testing data split
``` {r 3.Train and Test data split , echo=FALSE }
inTrain = createDataPartition(train_clean$classe, p = 0.8)[[1]]
training = train_clean[ inTrain,]
testing = train_clean[-inTrain,]
set.seed( 62433)
```
## Random Forest Model 
```{r 4. Creating Random Forest model, echo=FALSE}
set.seed(711)
train.rf <- randomForest(classe ~ ., data=training, importance=TRUE,ntree = 50 , mtry = 10 , proximity=TRUE)
print(train.rf)
predtrainrf <- predict(train.rf,newdata = testing)
confusionMatrix(testing$classe,predtrainrf)
```
## SVM Model
```{r 5. SVM Model , echo=FALSE}
modelsvm <- svm(classe ~ . , data = training)
predsvm <- predict(modelsvm,testing)
confusionMatrix(testing$classe,predsvm)
testpred <- predict(modelsvm,test_clean)
```
## PLOTTING of Predicted vs Actual Outcome
```{r 6. Plotting predicted vs actual , echo=FALSE}
qplot(classe,predsvm,colour=classe,data=testing)
```
## COnclusion 
### Based on the model results,Random Forest would be most appropriate model for this 
### dataset.